Towards a Practical Face Recognition System: \\ Robust Alignment and Illumination by Sparse Representation

### 2011-03-31 rebuttal.tex

   \title{\Large {\bf Manuscript  Revision Statement}}

   \author{{Andrew Wagner, John Wright, Arvind Ganesh, Zihan Zhou, Hossein Mobahi, and Yi Ma}}

   \date{}
   \maketitle

We again thank the reviewers and the area chair for their comments.
%I am mostly satisfied with the responses to the earlier review. The paper is
%in much better form now and more carefully written.
Changes have been made to the paper to address all of the changes requested by Reviewer 1,
including a more complete set of results in Table 2, Table 3 and Figure 11, and associated
discussions.  We give a quick overview of the changes.

{\bf
	Page 1 line 16; "there are many face recognition that fall roughly.."
	missing the word 'applications'
}

The typo on Page 1 is now amended.

{\bf
	In Table 1, the baseline results are shown with subscripts 'd' and 'm', i.e.
	using results of a detector, and manual registration. It seems like an obvious
	omission not to include results of recognition where input comes from the
	proposed registration algorithm itself. While the paper states that even
	manual registration works worse, it is not particularly convincing that manual
	registration is to be considered better than automatic registration - the
	human visual sensitivity is pretty bad for a problem like registration.
}

We ran the requested experiments.  We agree with the reviewer that automatic registration
can often out-perform manual registration.  The results back this up, and the text
is amended to be more clear on this point.  The new more complete benchmark is shown in
Table 2, and Figure 11.  All of the classical methods benefit from using our iterative
alignment procedure.  This is even true for LBP, which is known to have some robustness
to moderate misalignment.  The discussions have been updated to cover the new results.

{\bf
	In any case, the extent of manual input is unclear. It is said 'the input face
	is aligned to the training with manually selected outer eye corners'. If only
	2 points are manually selected, then what kind of registration can be done
	with just 2 corresponding points ? Surely not affine alignment. Many more are
	needed if it is to be done robustly. So it is not surprising that manual
	alignment is giving worse results than automatic. If my interpretation is
	wrong, is the entire affine registration being done manually ? If so, as
	mentioned above the accuracy of the human can be bad.
}

Similarity transformations are used for all of the experiments in this section, and
therefore two manually selected outer eye corners are sufficient.  We have
made this clear by adding a footnote.

{\bf
	Some comments on this would be useful to better understand the results shown in table 1 and 2.
}

The discussion of these results (now Table 2 and 3 in the new
draft) has been updated. For the single gallery image FERET
experiment (Table 3), since our algorithm is not designed for
such scenario, it fails when the illumination changes and other
variations seriously violate its assumption. We found that LBP
with manual alignment achieves the best performance in this
experiment.

### 2011-03-31 faces_system.tex
More numbers from Zihan

% Classical algorithms
To better examine the effectiveness of our iterative alignment
algorithm, we next compare our result to baseline
linear-projection-based algorithms, such as Nearest Neighbor
(NN), Nearest Subspace (NS) \cite{Lee2005-PAMI}, and Linear
Discriminant Analysis (LDA)
\cite{Belhumeur1997-PAMI}.\footnote{We do not list results on
PCA \cite{Turk1991-CVPR} as its performance is always below
that of Nearest Subspace.} Since these algorithms assume
pixel-accurate alignment, they are not expected to work well if
the test image is not well aligned with the training. In
Table~\ref{tab:MultiPIE-recognition}, we report the results of
these classical algorithms with three types of testing image
alignment: 1.\ alignment from the Viola and Jones' detector,
2.\ alignment via manually selected outer eye corners,
\footnote{Two manually clicked points are sufficient to define
a similarity transformation.  All of the experiments in this section are carried out with
similarity transformations.}
 and 3.\
the output of our iterative alignment algorithm. The
performance drop of the LDA algorithm on Multi-PIE reported
here seems to agree with that reported already in
\cite{Gross2008-FGR}.  All of the classical algorithms benefit
greatly from being paired with our iterative alignment
algorithm.

@@

@@

In addition, although our algorithm is not designed for
recognition when there is only a single gallery image per user,
we compare its performance with LBP within this setting for
completeness. For this experiment, we use the FERET dataset
\cite{phillips1998feret}, which contains five standard
partitions: `fa' is the gallery containing 1196 frontal images
of 1196 subjects, and `fb', `fc', `dup1' and `dup2' are four
sets of probe images. The testing sets differ from the training
in facial expression (`fb'), illumination (`fc'), aging (`dup1'
) and long aging (`dup2'). In fact, except for `fb', we notice
significant changes of illumination in all the other three test
sets. For the training, we again crop and normalize the face
region from each original image to an $80\times 60$ window
using manually marked eye coordinates \cite{Deng2010-PR}. In
Table~\ref{tab:FERET-recognition}, we report the performance of
our algorithm on the four test sets, with input directly
obtained from the Viola and Jones' detector, and the
performance of LBP with input obtained from both face detector
and manual alignment using the eye coordinates.

\begin{table}
\caption{Recognition rates on single gallery image FERET
dataset} \vspace{-.1in} \centerline{
\begin{tabular}{|c|c|c|c|c| }
\hline
Recognition rate & fb & fc & dup1 & dup2 \\
\hline
\hline
LBP$_d$ (LBP$_m$) & 54.8 (96.6)\% & 10.3 (58.8)\% & 29.8 (71.6)\% & 19.8 (61.5)\% \\
\hline
{Algorithm 1} & 95.2\% & 28.4\% & 46.1\% & 20.3\% \\
\hline
\end{tabular}
\label{tab:FERET-recognition} } \vspace{-.4in}
\end{table}

As expected, our algorithm does not perform well except for
`fb', in which the illumination is similar to the training and
the mere variation in facial expression is handled well by the sparse error
model. For the other three test sets, our algorithm fails because the illumination
changes and other variations seriously violate the assumptions of our method.
On the other hand, while LBP achieves the best recognition rates given
manually aligned training and testing samples, its performance
degrades drastically when the input is obtained directly from
the face detector. It is also worth noting that similar poor
performance of LBP, as well as other descriptors, has been
observed on the Labeled Face in the Wild (LFW) database, where
the training is uncontrolled and limited and the input is
directly obtained from the face detector \cite{Wolf2008-ECCV}.

>>>>

% Classical algorithms
To better examine the effectiveness of our iterative alignment
algorithm, we next compare our result to baseline
linear-projection-based algorithms, such as Nearest Neighbor
(NN), Nearest Subspace (NS) \cite{Lee2005-PAMI}, and Linear
Discriminant Analysis (LDA)
\cite{Belhumeur1997-PAMI}.\footnote{We do not list results on
PCA \cite{Turk1991-CVPR} as its performance is always below
that of Nearest Subspace.} Since these algorithms assume
pixel-accurate alignment, they are not expected to work well if
the test image is not well aligned with the training. In
Table~\ref{tab:MultiPIE-recognition}, we report the results of
these classical algorithms with three types of testing image
alignment: 1.\ alignment from the Viola and Jones' detector,
2.\ alignment via manually selected outer eye
corners,\footnote{Two manually clicked points are sufficient to
define a similarity transformation. All of the experiments in
this section are carried out with similarity transformations.}
and 3.\ the output of our iterative alignment algorithm. The
performance drop of the LDA algorithm on Multi-PIE reported
here seems to agree with that reported already in
\cite{Gross2008-FGR}.  All of the classical algorithms benefit
greatly from being paired with our iterative alignment
algorithm.

@@

%\begin{table}
%\caption{Recognition rates on the Multi-PIE database for
%different pairings of alignment and recognition stages.}
%\vspace{-.2in} \centerline{
%\begin{tabular}{|c|c|c|c|c| } \hline
%Recognition rate & Session 2 & Session 3 & Session 4  \\
%\hline
%LDA($d/m/i$) & 5.1 / 49.4 / 91.0\% & 5.9 / 44.3 / 89.9\% & 4.3 / 47.9 / 88.1\% \\
%\hline
%NN($d/m/i$) & 26.4 / 67.3 / 73.5\% & 24.7 / 66.2 / 69.6\% & 21.9 / 62.8 / 69.3\% \\
%\hline
%NS($d/m/i$) & 30.8 / 77.6 / 84.5\% & 29.4 / 74.3 / 82.3\% & 24.6 / 73.4 / 81.4\% \\
%\hline
%LBP($d/m/i$) & 39.9 / 93.3 / {\bf 95.2}\% & 38.1 / 91.2 / {\bf 94.7}\% & 33.9 / 92.9 / {\bf 93.5}\% \\
%\hline
%Algorithm 1 & 93.9\% & 93.8\% & 92.3\% \\
%\hline
%\end{tabular}
%\label{tab:MultiPIE-recognition} }\vspace{-.4in}
%\end{table}

@@

In addition, although our algorithm is not designed for
recognition when there is only a single gallery image per user,
we compare its performance with LBP within this setting for
completeness. For this experiment, we use the FERET dataset
\cite{phillips1998feret}, which contains five standard
partitions: `fa' is the gallery containing 1196 frontal images
of 1196 subjects, and `fb', `fc', `dup1' and `dup2' are four
sets of probe images. The testing sets differ from the training
in facial expression (`fb'), illumination (`fc'), aging (`dup1'
) and long aging (`dup2'). In fact, except for `fb', we notice
significant changes of illumination in all the other three test
sets. For the training, we again crop and normalize the face
region from each original image to an $80\times 60$ window
using manually marked eye coordinates \cite{Deng2010-PR}. In
Table~\ref{tab:FERET-recognition}, we report the performance of
our algorithm on the four test sets, with input directly
obtained from the Viola and Jones' detector. We also report the
performance of LBP with the same three types of input as before
in Table~\ref{tab:FERET-recognition}, where we use letters
``$d$'', ``$m$'' and ``$i$'' to indicate face detector, manual
alignment and our iterative alignment algorithm, respectively.

\begin{table}
\caption{Recognition rates on single gallery image FERET
dataset} \vspace{-.1in} \centerline{
\begin{tabular}{|c|c|c|c|c| }
\hline
Recognition rate & fb & fc & dup1 & dup2 \\
\hline
\hline
LBP($d/m/i$) & 54.8 / {\bf 96.6} / 94.5 \% & 10.3 / {\bf 58.8} / 42.8\% & 29.8 / {\bf 71.6} / 46.5\% & 19.8 / {\bf 61.5} / 21.1\% \\
\hline
{Algorithm 1} & 95.2\% & 28.4\% & 46.1\% & 20.3\% \\
\hline
\end{tabular}
\label{tab:FERET-recognition} } \vspace{-.4in}
\end{table}

As expected, our algorithm does not perform well except for
`fb', in which the illumination is similar to the training and
the mere variation in facial expression is handled well by the
sparse error model. For the other three test sets, our
algorithm fails because the illumination changes and other
variations seriously violate the assumptions of our method.
This also explains why LBP performs worse with our iterative
alignment algorithm, compared to manual alignment. On the other
hand, while LBP achieves the best recognition rates given
manually aligned training and testing samples, its performance
degrades drastically when the input is obtained directly from
the face detector. It is also worth noting that similar poor
performance of LBP, as well as other descriptors, has been
observed on the Labeled Face in the Wild (LFW) database, where
the training is uncontrolled and limited and the input is
directly obtained from the face detector \cite{Wolf2008-ECCV}.
